Introduction

### 2020-10-14 text/introduction.Rmd
minor edits to introduction

Recent advances in technology have greatly expanded our view into the mechanisms
of cancer at the systems biology level. Next-generation sequencing technologies
have made it possible to assay germline and somatic mutations, expression
profiles, DNA methylation, and copy number variations from the same tissue. To
take advantage of these tools, several large consortia, including the cancer
genome atlas (TCGA) and pan-cancer analysis of whole genomes (PCAWG) sequenced
large numbers of tumors multiple assays to analyze together, with the goal of
integration and increased understanding of the mechanisms of cancer.

@@

More recently, pan-cancer mutations and other data have been looked at as
disruptions of normal pathway activity. @PCAWG-PATH carried out exhaustive
analysis of pathway enrichment using whole genome data. Network approaches
address the sparseness of mutations, allowing genes to be influenced by
mutations in nearest network neighbors [@hotnet2;@hornNetSigNetworkbasedDiscovery2018;
@Iranzo-Network]. These methods are powerful especially when applied to smaller
numbers of tumors. One of the challenges in the field remains that there may be
thousands of moderate effect genes that occur at such low frequency--given
tissue diversity and low sample number--that they are impossible to detect using
positive selection theory. Some groups have attempted to address this challenge
by applying machine learning to cancer data to discover groups of
functionally related genes as they interact with larger pathways and networks
[@kimPancancerAnalysisSomatic2018;
@mourikis-ml-pathways; @moonlight]. These studies have proven very effective at
highlighting fundamental disease phenotypes at the pathway level across cancers
with different origins at the cellular and tissue level.

>>>>

Advances in technology have greatly expanded our view into the mechanisms
of cancer at the systems biology level. Next-generation sequencing technologies
have made it possible to study germline and somatic mutations, expression
profiles, DNA methylation, and copy number variations from the same tissue. To
take advantage of these tools, several large consortia, including the cancer
genome atlas (TCGA) and pan-cancer analysis of whole genomes (PCAWG) sequenced
large numbers of tumors multiple assays to analyze together, with the goal of
integration and increased understanding of the mechanisms of cancer.

@@

More recently, pan-cancer mutations and other data have been looked at as
disruptions of normal pathway activity. @PCAWG-PATH carried out exhaustive
analysis of pathway enrichment using whole genome data. Network approaches
address the sparseness of mutations, allowing genes to be influenced by
mutations in nearest network neighbors [@hotnet2;@hornNetSigNetworkbasedDiscovery2018;
@Iranzo-Network]. These methods are powerful especially when applied to smaller
numbers of tumors. One of the challenges in the field remains that there may be
thousands of moderate effect genes that occur at such low frequency that they
are impossible to detect using positive selection theory. Some researchers have
attempted to address this challenge by applying machine learning to cancer data
to discover groups of functionally related genes as they interact with larger
pathways and networks [@kimPancancerAnalysisSomatic2018;
@mourikis-ml-pathways; @moonlight]. These studies have proven very effective at
highlighting fundamental disease phenotypes at the pathway level across cancers
with different origins at the cellular and tissue level.
